Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating subgenome assignments in maize

### 2017-08-11 R/SubGenome Analysis.Rmd
Added titles to plots

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
library(VennDiagram)
```

@@

# SubGenomes
***
In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. (TODO)

@@

## Compared to Paper
How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2? Items in "Paper" only are missed subgenome assignments.  Items in "Test" only are false assignments.  Items in center are correct. (this is an unfair comparison if I sorted by hand)

@@

grid.newpage()
draw.pairwise.venn(
  area1 = subgenome.truth %>% nrow(), 
  area2 = subgenome.chromosomes.test %>% nrow(), 
  cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
  category = c("Paper", "Test"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

Genes placed in the "correct" subgenome (TODO)

@@

# GOTerms
***
GO Term types by EVCode

```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
Type <- c("EV-EXP","EV-COMP")
Count <- c(goAnnotations.mutated %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
      goAnnotations.mutated %>% filter(grepl("EV-COMP", EVCode)) %>% nrow())
df = data.frame(Type,Count)
ggplot(data=df, aes(x=Type, y=Count)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_text(aes(label=Count), vjust=-0.3, size=3.5)
```

GO Term by subgenome (All EV-Codes)

```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# All EV-Codes, Unique GO Terms
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% select("GO Term") %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% select("GO Term") %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% select("GO Term"), goAnnotations.sub2 %>% select("GO Term")) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)

#EXP only, Unique GO Terms
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
                         ) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

GOTerm assignments to genes by subgenome (All EV-Codes)
(broken - doesn't take into account homeologous genes - try replacing equivalent genes from homeolog list)

```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# All EV-Codes, all assignments
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% nrow(),
  area2 = goAnnotations.sub2 %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term"), goAnnotations.sub2 %>% select(gene2, "GO Term")) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)

#EXP only, all assignments
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term"), 
                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term")
                         ) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
library(VennDiagram)
library(gridExtra)
```

@@

# SubGenomes
***
In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. 
**(TODO)**

@@

## Compared to Paper
How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2? Items in "Paper" only are missed subgenome assignments.  Items in "Greedy" only are false assignments.  Items in center are correct. (this is an unfair comparison if I sorted by hand)

@@

grid.newpage()
g = draw.pairwise.venn(
  area1 = subgenome.truth %>% nrow(), 
  area2 = subgenome.chromosomes.test %>% nrow(), 
  cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
  category = c("Paper", "Greedy"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="Overlap of Subgenome Assignments vs. \"Truth\"")
```

Genes placed in the "correct" subgenome

**(TODO)**

@@

# GOTerms
***

```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
Type <- c("EV-EXP","EV-COMP")
Count <- c(goAnnotations.mutated %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
      goAnnotations.mutated %>% filter(grepl("EV-COMP", EVCode)) %>% nrow())
df = data.frame(Type,Count)
ggplot(data=df, aes(x=Type, y=Count)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_text(aes(label=Count), vjust=-0.3, size=3.5
) + labs(
  title="GOTerms by EVCode"
)
```

***

GO Terms by subgenome.  If the GO Term is assigned to any gene in the subgenome, it is counted once for that subgenome.  This is a presence/absence test.

```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# All EV-Codes, Unique GO Terms
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% select("GO Term") %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% select("GO Term") %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% select("GO Term"), goAnnotations.sub2 %>% select("GO Term")) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term by subgenome (All EV-Codes)")

#EXP only, Unique GO Terms
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
                         ) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term by subgenome (EV-EXP only)")
```

***

GOTerm assignments to genes by subgenome, which takes into account the gene-GOTerm pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.

(broken - doesn't take into account homeologous genes - try replacing equivalent genes from homeolog list)

```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# All EV-Codes, all assignments
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% nrow(),
  area2 = goAnnotations.sub2 %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term"), goAnnotations.sub2 %>% select(gene2, "GO Term")) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term assignments by subgenome (EV-EXP only)")

#EXP only, all assignments
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term"), 
                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term")
                         ) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term assignments by subgenome (EV-EXP only)")
```
